Analysing the Performance of Shallow-Water Models with the Roofline Model

### 2016-01-19 docs/papers/roofline_model/gocean_roofline.tex
roofline: add more results to doc

@@

>>>>

\section{Introduction}

We want to use the Roofline Model to establish the efficiency (or
otherwise) of two shallow-water benchmark codes that we have used in
the GOcean project.

\section{Hardware Details}

CPU name:       Intel(R) Xeon(R) CPU E5-1620 v2 @ 3.70GHz
CPU type:       Intel Xeon IvyBridge EN/EP/EX processor

Level 1 32 kB
Level 2 256 kB
Level 3 10 MB

\section{Hardware Performance Counters}

The likwid tool (\url{https://github.com/RRZE-HPC/likwid}) gives
access to Hardware Performance Counters (HWPCs) on Intel and AMD CPUs.
However, there are issues with the accuracy of these counters,
especially for FLOP-related measurements on Ivybridge architectures. A
tool is included in the likwid distribution that tests the accuracy of
HWPC-derived data by running micro-benchmarks for which the FLOP count
and memory traffic are known:
\url{https://code.google.com/p/likwid/wiki/IvyBridgeVerification}.
The results of doing this on an Intel Xeon E5-1620 v2 CPU are shown in
Figures~\ref{FIG_flops_dp_stream_test}
and~\ref{FIG_flops_dp_triad_test}. It appears that once the memory
footprint spills out of L1 cache (which all of our nemolite2d
benchmark configurations do), the performance counter results
significantly overestimate the FLOPs performed.

\begin{figure}
\centering
\includegraphics[width=120mm]{FLOPS_DP_stream_ivybridgeEP.pdf}
\caption{Results of the likwid accuracy test for the FLOPS\_DP group
  and the stream benchmark on an Intel Xeon E5-1620 v2.}
\label{FIG_flops_dp_stream_test}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=120mm]{FLOPS_DP_triad_ivybridgeEP.pdf}
\caption{Results of the likwid accuracy test for the FLOPS\_DP group
  and the triad benchmark on an Intel Xeon E5-1620 v2.}
\label{FIG_flops_dp_triad_test}
\end{figure}

We therefore need some other way of measuring FLOPS. We could use
likwid on an older Intel CPU (to which we have root access) or try
another tool on the IvybridgeEP system. Other possible tools include
Intel's VTune or OProfile.

@@

\begin{table}
\begin{tabular}{l|l|l|l|l}
\hline
\hline
               & \multicolumn{4}{c}{Problem size} \\
               & 32     &    64        &    128       &   256 \\
\hline
Footprint (MB) & 0.1 & 0.5          & 2.3          & 9.2 \\
\hline
Runtime (RDTSC) [s]   & 0.144868 &  0.4485108   &   1.697142   &   6.446171  \\
Runtime unhalted [s]  & 1.199456e-01 & 4.475135e-01 & 1.750448e+00 &   6.734944e+00 \\
CPI                   & 6.431867e-01 & 6.423220e-01 & 6.503796e-01 &   6.348858e-01 \\
\hline
MFLOPS           & 3.112780e+03 & 3.877720e+03  & 4.163209e+03 & 4.304728e+03 \\
AVX MFLOPS       & 2.840625e+03 & 3.660086e+03  & 3.993397e+03 & 4.164647e+03 \\
Packed MUOPS/s   & 7.423377e+02 & 9.584935e+02  & 1.046781e+03 & 1.092280e+03 \\
Scalar MUOPS/s   & 2.077927e+02 & 1.306897e+02  & 7.294811e+01 & 3.784506e+01 \\
\hline
L1 DTLB load misses          &      2007     &      17015   &    1317912   &    6748712   \\
L1 DTLB load miss rate       &  2.949686e-06 &  6.607264e-06& 1.326630e-04 & 1.723565e-04 \\
L1 DTLB load miss durn [Cyc] &  3.970154e+01 &  3.080153e+01& 2.101039e+01 & 2.013584e+01 \\
L1 DTLB store misses         &       77      &       16     &     87257    &    401036    \\
L1 DTLB store miss rate      &  1.131668e-07 &  6.213119e-09& 8.783423e-06 & 1.024213e-05 \\
L1 DTLB store miss durn [Cyc]&  5.633766e+01 &       29     & 2.151055e+01 & 2.669793e+01 \\
\hline
L2D load bandwidth [MB/s]  &  3.811099e+03 &  5.030691e+03 & 1.080936e+04 & 1.446214e+04 \\
L2D load data volume [GB]  &   0.541220288 &   2.34655488  &  18.1215488  &   92.177536  \\
L2D evict bandwidth [MB/s] &  3.131102e+02 &  3.094459e+02 & 4.501487e+02 & 5.065637e+02 \\
L2D evict data volume [GB] &   0.04446528  &   0.144340352 &  0.75465984  &  3.22869248  \\
L2 bandwidth [MB/s]        &  4.600834e+03 &  5.493404e+03 & 1.130282e+04 & 1.498363e+04 \\
L2 data volume [GB]        &   0.653371776 &   2.562386304 & 18.948807808 & 95.501414592 \\
\hline
L3 load bandwidth [MB/s]  &  2.434255e+03 & 4.626177e+03 & 4.511686e+03 & 4.607617e+03 \\
L3 load data volume [GB]  &   0.371968512 &  2.07489024  &   7.6168768  &   29.692832  \\
L3 evict bandwidth [MB/s] &  5.652167e+02 & 4.828169e+02 & 2.771148e+02 & 2.722823e+02 \\
L3 evict data volume [GB] &   0.086368448 &  0.216548608 &  0.467840512 &  1.75466688  \\
L3 bandwidth [MB/s]       &  2.999472e+03 & 5.108994e+03 & 4.788801e+03 & 4.879899e+03 \\
L3 data volume [GB]       &   0.45833696  &  2.291438848 &  8.084717312 &  31.44749888 \\
\hline
\hline
\end{tabular}
\caption{Hardware performance counter data for the u-Momentum kernel on the Intel E5-1620 v2 processor.}
\label{TAB_HWPC_MOM_U}
\end{table}
